---
format:
  html:
    page-layout: article
title: "ational Influenza Genomic Surveillance Report"
author:
  - name: Jane Doe
    email: jane.doe@mail.com
    corresponding: true
  - name: John Doe
date: today
execute:
  echo: false
  warning: false
  message: false
  error: false
params:
  meta_file: "gisaid_epiflu_isolates.xls"
  tree_file: "tree.nwk"
  country: "United States"
  
---

**Country:**  
**Influenza Season:** (e.g., 2025 Southern Hemisphere)  
**Reporting Period:** (Start date – End date)  
**Prepared by:** National Influenza Center / Genomic Surveillance Unit  
**Date of Report:**  

```{r}
# install.packages("tidyverse")
# install.packages("BiocManager")
# BiocManager::install("ggtree")
# install.packages('lubridate')
# install.packages("readxl")
# install.packages("sf")
# install.packages("rnaturalearth")
# install.packages("rnaturalearthdata")
# install.packages("countrycode")
# install.packages("scales")
# install.packages("scatterpie")
# install.packages("phangorn")
# install.packages("patchwork")
```

```{r}
library(tidyverse)
library(ggtree)
library(knitr)
library(lubridate)
library(readxl)

library(sf)
library(rnaturalearth)
library(countrycode)
library(scales)
library(scatterpie)
library(phangorn)
library(patchwork)
```

```{r}
df <- read_excel(params$meta_file) %>%
  mutate(
    sid     = str_split_fixed(`HA Segment_Id`, "\\|", 2)[, 1],
    date    = ymd(Collection_Date),
    month = floor_date(date, "month"),
    country = trimws(str_split_fixed(Location, "/", 4)[, 2]),
    clade   = apply(str_split_fixed(Clade, "\\.", 6)[, 1:3], 1, paste, collapse = ".")
  ) %>%
  select(sid, date, month, country, clade) %>%
  mutate(
    clade = as.factor(clade)
  )

# ---- consistent clade colors for the whole document ----
clade_levels  <- levels(droplevels(df$clade))
clade_palette <- setNames(hue_pal()(length(clade_levels)), clade_levels)

df <- df |> 
  mutate(
    clade = factor(clade, levels = clade_levels)
  )
```

## 1. Executive Summary

This report summarizes influenza virus genomic surveillance conducted during the **[season]** influenza season in **[country]**. Surveillance integrates routine virological testing with whole-genome sequencing to monitor circulating influenza strains, genetic diversity, antiviral resistance markers, and phylogenetic relationships.

Key findings:
- Total specimens tested: **[N]**
- Influenza-positive specimens: **[N]**
- Viruses sequenced: **[N]**
- Dominant clades/subclades: **[e.g., A(H1N1)pdm09 clade 6B.1A.5a]**
- Antiviral resistance marker H275Y detected in **[N]** viruses

---

## 2. Surveillance System and Sample Collection

### 2.1 Surveillance Design

Specimens were collected through:
- ☐ Sentinel outpatient influenza-like illness (ILI) surveillance  
- ☐ Severe acute respiratory infection (SARI) surveillance  
- ☐ Hospital-based surveillance  
- ☐ Outbreak investigations  
- ☐ Other: **[specify]**

### 2.2 Sample Collection Method

- **Specimen type:** (e.g., nasopharyngeal swab, oropharyngeal swab)
- **Collection medium:** (e.g., viral transport medium)
- **Testing laboratories:** **[list]**
- **Age groups included:** **[e.g., all ages / ≥6 months]**
- **Geographic coverage:** **[national / subnational regions]**

### 2.3 Sample Collection Timeline
```{r}
#| fig-cap: "Figure 1. Sample collection times of reported clades."

df |>
  drop_na(date) |>
  count(clade, month, name = "n") |>
  complete(
    clade,
    month = seq(min(month), max(month), by = "month"),
    fill = list(n = 0)
  ) |>
  arrange(month) |>
  ggplot(aes(month, n, fill = clade)) +
  geom_area() +
  scale_fill_manual(values = clade_palette, drop = FALSE) +
  theme_bw() +
  labs(x = "Collection Date", y = "Number Reported", fill = "Clade")
```

### 2.4 Sample Collection Locations
```{r}
#| fig-cap: "Figure 2. Geographic distribution of reported clades."

# World polygons (sf)
world <- rnaturalearth::ne_countries(scale = "medium", returnclass = "sf") |>
  select(iso_a3, name_long, geometry)

# ---- clade composition per country (wide) ----
df_country_wide <- df |>
  drop_na(country, clade) |>
  mutate(
    iso_a3 = countrycode(country, origin = "country.name", destination = "iso3c"),
    clade  = factor(clade, levels = clade_levels)
  ) |>
  drop_na(iso_a3) |>
  count(iso_a3, clade, name = "n") |>
  tidyr::pivot_wider(names_from = clade, values_from = n, values_fill = 0) |>
  mutate(total_n = rowSums(across(all_of(clade_levels))))

# ---- centroids for each country polygon ----
world_pts <- world |>
  st_centroid(of_largest_polygon = TRUE) |>
  mutate(
    lon = st_coordinates(geometry)[, 1],
    lat = st_coordinates(geometry)[, 2]
  ) |>
  st_drop_geometry() |>
  select(iso_a3, lon, lat)

# ---- join pies to centroids ----
pies <- df_country_wide |>
  inner_join(world_pts, by = "iso_a3")

# ---- plot: base map + pie charts ----
world |>
  ggplot() +
  geom_sf(fill = "grey95", color = "white", linewidth = 0.1) +
  scatterpie::geom_scatterpie(
    data = pies,
    aes(x = lon, y = lat, r = sqrt(total_n)),   # size pies by sample count
    cols = clade_levels,
    color = NA
  ) +
  coord_sf(expand = FALSE) +
  scale_fill_manual(values = clade_palette, drop = FALSE) +
  theme_bw() +
  labs(fill = "Clade") +
  theme(
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    panel.grid = element_blank()
  )
```

## 3. Virological Screening Results

### 3.1 Total Specimens Tested

| Metric | Count |
|------|------:|
| Total specimens tested | |
| Influenza-positive specimens | |
| Influenza-negative specimens | |

### 3.2 Influenza Type and Subtype Distribution

| Virus Type / Subtype | Count | Percentage (%) |
|---------------------|------:|---------------:|
| Influenza A (total) | | |
| └─ A(H1N1)pdm09 | | |
| └─ A(H3N2) | | |
| Influenza B (total) | | |
| └─ B/Victoria lineage | | |
| └─ B/Yamagata lineage | | |
| **Total influenza positive** | | 100 |

---

## 4. Sequencing Overview

### 4.1 Sequencing Selection Criteria

Specimens were selected for sequencing based on:
- ☐ Ct value ≤ **[threshold]**
- ☐ Geographic representation
- ☐ Temporal distribution across season
- ☐ Age group or clinical severity
- ☐ Other: **[specify]**

### 4.2 Sequencing Output and Quality

| Metric | Count |
|------|------:|
| Influenza-positive specimens eligible for sequencing | |
| Specimens submitted for sequencing | |
| Successfully sequenced genomes | |
| Genomes passing QC thresholds | |
| Genomes failing QC | |

**Passing rate:** **[X%]** (passing QC / sequenced)

QC thresholds included:
- Minimum genome coverage: **[e.g., ≥90%]**
- Maximum ambiguous bases (Ns): **[e.g., ≤5%]**

---

## 5. Clade and Subclade Distribution (Nextclade)

Clade assignment was performed using **Nextclade** (version **[X]**, dataset **[Y]**).

### 5.1 Influenza A(H1N1)pdm09

| Clade / Subclade | Count | Percentage (%) |
|------------------|------:|---------------:|
| | | |
| | | |
| **Total** | | 100 |

### 5.2 Influenza A(H3N2)

| Clade / Subclade | Count | Percentage (%) |
|------------------|------:|---------------:|
| | | |
| | | |
| **Total** | | 100 |

### 5.3 Influenza B

| Lineage / Clade | Count | Percentage (%) |
|-----------------|------:|---------------:|
| | | |
| | | |
| **Total** | | 100 |

---

## 6. Antiviral Resistance Markers

### 6.1 Neuraminidase H275Y (A(H1N1)pdm09)

| Metric | Count |
|------|------:|
| A(H1N1)pdm09 genomes analyzed | |
| Viruses with H275Y mutation | |
| Percentage with H275Y | |

**Interpretation:**  
The H275Y substitution is associated with reduced susceptibility to oseltamivir.

---

## 7. Phylogenetic Analysis

### 7.1 Methods

- **Alignment:** **[e.g., MAFFT]**
- **Phylogenetic inference:** **[e.g., IQ-TREE / FastTree]**
- **Reference strains:** WHO vaccine reference strains for the **[season]** season
- **Gene(s) analyzed:** ☐ HA ☐ NA ☐ Whole genome

### 7.2 Phylogenetic Tree Summary

- Circulating viruses clustered primarily with **[reference clade/strain]**
- No major divergent clusters were observed ☐ / Divergent clusters were observed ☐
- Genetic diversity was **[low / moderate / high]**

### 7.3 Phylogenetic Tree

# Phylogenetic tree
```{r}
#| fig-cap: "Figure 3. Phylogenetic tree with clade annotation."

tree <- read.tree(params$tree_file) |>
  midpoint()

# ---- 1) Build tree plot and extract *plotted* tip order ----
p_tree <- ggtree(tree)

tip_order <- p_tree$data |>
  dplyr::filter(isTip) |>
  dplyr::arrange(y) |>
  dplyr::pull(label)

df_tree <- df |>
  dplyr::filter(sid %in% tip_order) |>
  dplyr::mutate(sid = factor(sid, levels = tip_order))

# ---- 2) Clade tile (1 column) ----
p_clade <- ggplot(df_tree, aes(x = "Clade", y = sid, fill = clade)) +
  geom_tile(width = 0.95, height = 0.95) +
  scale_y_discrete(limits = tip_order, expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0)) +
  labs(x = NULL, y = NULL) +
  theme_minimal(base_size = 10) +
  theme(
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    panel.grid = element_blank(),
    legend.position = "right",
    plot.margin = margin(5.5, 0, 5.5, 5.5)
  )+
  scale_fill_manual(values = clade_palette, drop = FALSE)+
  labs(fill="Cade")


# ---- 4) Combine with patchwork ----
final <- p_tree + p_clade +
  plot_layout(widths = c(4, 0.5))

final
```

---

## 8. Data Sharing

- Sequences submitted to: ☐ GISAID ☐ GenBank
- Number of sequences shared: **[N]**
- Accession IDs: **[range or link]**

---

## 9. Limitations

- Sequencing coverage may not be fully representative of all regions.
- Selection bias due to Ct-based sequencing thresholds.
- Turnaround time varied during peak transmission periods.

---

## 10. Conclusions and Public Health Implications

Genomic surveillance during the **[season]** influenza season demonstrated **[key findings]**, supporting national and global influenza risk assessment and vaccine strain selection processes.

---

## 11. Acknowledgements

We acknowledge sentinel sites, laboratories, clinicians, and data contributors supporting influenza surveillance in **[country]**.

---
